---
title: "COVID-19 age-specific reporting rates in Austria"
description: |
  Exploration of the temporal evolution of the age distribution of 
  COVID-19 reporting rates in Austria.
author:
- name: Fabian Valka
  affiliation: vektorraum
  affiliation_url: https://www.vektorraum.com/
date: "`r Sys.Date()`"
citation_url: https://fvalka.github.io/covid-austria-age-heterogenity/
output: distill::distill_article
---

<!-- Matomo -->
<script type="text/javascript">
  var _paq = window._paq || [];
  /* tracker methods like "setCustomDimension" should be called before "trackPageView" */
  _paq.push(["setDoNotTrack", true]);
  _paq.push(["disableCookies"]);
  _paq.push(['trackPageView']);
  _paq.push(['enableLinkTracking']);
  (function() {
    var u="//analytics.covid19-r.com/";
    _paq.push(['setTrackerUrl', u+'matomo.php']);
    _paq.push(['setSiteId', '5']);
    var d=document, g=d.createElement('script'), s=d.getElementsByTagName('script')[0];
    g.type='text/javascript'; g.async=true; g.defer=true; g.src=u+'matomo.js'; s.parentNode.insertBefore(g,s);
  })();
</script>
<!-- End Matomo Code -->

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(RcppRoll)
library(forecast)

source(here("R/functions/fix-negative-incidence.R"))
source(here("R/functions/austrian-demography.R"))
source(here("R/pipeline/convert-age-stratified-time-series.R"))

convert_age_stratified_time_series()

population_stratified <- gather(
  as_tibble(
    as.list(c(
      "<5"=demo_age_group_total(0:4),
      "5-14"=demo_age_group_total(5:14),
      "15-24"=demo_age_group_total(15:24),
      "25-34"=demo_age_group_total(25:34),
      "35-44"=demo_age_group_total(35:44),
      "45-54"=demo_age_group_total(45:54),
      "55-64"=demo_age_group_total(55:64),
      "65-74"=demo_age_group_total(65:74),
      "75-84"=demo_age_group_total(75:84),
      ">84"=demo_age_group_total(85:100),
      ">24"=demo_age_group_total(25:100)))), "age_group", "pop")

plot_order <- c("<5", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75-84", ">84")

data <- read.csv(here("data/age-stratified-incidence-austria.csv"))

data <- data %>%
  group_by(age_group) %>%
  mutate(incidence=fix_negative_incidence(incidence)) %>% 
  mutate(incidence_7day = roll_meanr(incidence, 7)) %>% 
  left_join(population_stratified) %>%
  mutate(incidence_7day_per_capita = incidence_7day/pop*10e5) %>%
  mutate(incidence_per_capita = incidence/pop*10e5) %>%
  filter(row_number() > 7 & row_number() < n()) %>%
  mutate(date=ymd(date))

rel_incidence <- data %>%
  mutate(incidence_per_capita_rel = incidence_7day_per_capita/max(incidence_7day_per_capita)) %>%
  filter(n() - row_number() <= 45)

pal_grad3 <- c("#10547E", "#82BC24", "#FFCE47", "#FF0A12")

last_update <- max(data$date)

```

# New cases per 100k people in age group

The reporting rate for each age-group is calculated by dividing the 
number of reported COVID-19 cases in the age-group by the number of people
in that specific age-group. 

Last update: `r last_update`

### Heatmap, 7-day moving average

```{r, echo=FALSE, fig.fullwidth=TRUE, layout="l-page"}
age_density_plot <- data  %>%
  ggplot() +
  geom_raster(aes(x=date, y=age_group, fill=incidence_7day_per_capita, label=incidence_7day, label2=incidence)) +
  scale_fill_gradientn(colours = colorRampPalette(pal_grad3,space="rgb", bias = 1.7)(1000),
                       limits=c(0,200),
                       n.breaks=5,
                       guide = guide_colourbar(barwidth = 25,
                                               title.position = "top",
                                               title.hjust = 1)
  ) +
  scale_y_discrete(limits = plot_order) +
  scale_x_date(date_minor_breaks = "1 week") + 
  theme_pubclean() +
  labs(fill="Cases\nper 100k", x=NULL, y="Age Group") +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5, face = "bold", size=21, 
                                  margin=margin(12,12,15,12)))

plotly::ggplotly(age_density_plot, options=list(digits=1))

```

### Plot

```{r, echo=FALSE, fig.fullwidth=TRUE, layout="l-page"}
(data %>%
   mutate(incidence_per_capita_rel = incidence_7day_per_capita/max(incidence_7day_per_capita)) %>%
   ggplot() +
   geom_line(aes(x=date, y=incidence_7day_per_capita, color=age_group)) +
   scale_color_viridis_d(limits=plot_order, option = "plasma", direction=-1) +
   theme_pubclean() +
   labs(color="Age group") +
   xlab("Date") + 
   ylab("New cases per 100k people")) %>%
  plotly::ggplotly()
```

# Proportions of the age-groups

Proportions of the cases per 100k in each age-group to the sum of 
cases per 100k in all age-groups. Both are smoothed using the 7-day moving average. 

This visualization shows how over represented each age group is compared to
the other age-groups, adjusted for the different population sizes in each age group. 

### Heatmap

```{r, echo=FALSE, fig.fullwidth=TRUE, layout="l-page"}

total_incidence <- data %>% 
  group_by(date) %>% 
  summarise(total_incidence_7day_per_capita = sum(incidence_7day_per_capita))

(data  %>%
  left_join(total_incidence) %>%
  mutate(relative_incidence = incidence_7day_per_capita/total_incidence_7day_per_capita) %>%
  ggplot() +
  geom_raster(aes(x=date, y=age_group, fill=relative_incidence)) +
  scale_fill_gradientn(colours = colorRampPalette(pal_grad3,space="rgb")(1000),
                       limits=c(0, 0.7),
                       n.breaks=5,
                       guide = guide_colourbar(barwidth = 25,
                                               title.position = "top",
                                               title.hjust = 1)
  ) +
  scale_y_discrete(limits = plot_order) +
  theme_pubclean() +
  labs(fill="Proportion", x=NULL, y="Age Group") +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5, face = "bold", size=21, margin=margin(12,12,15,12)))) %>%
  plotly::ggplotly()
```

# Methods and sources

#### Reported cases by age group

The cumulative number of reported cases is published in the Austrian 
open government data portal:<br>
[BMSGPK, Österreichisches COVID-19 Open Data Informationsportal (https://www.data.gv.at/covid-19).](https://www.data.gv.at/covid-19)

Based upon a 
[data lake with hourly snapshots, by Attila Kerekes](https://github.com/keriati/austria-covid-data),
a time series is reconstructed from the hourly snapshots of the cumulative data by 
calculating the difference between the last snapshot of the day and the last 
snapshot of the previous day. 

If a day has a negative number of new cases this is corrected by applying the 
negative number of cases to the previous days such that each day has a non-negative
notification rate. 

#### Austrian population by age-group

Eurostat, The Statistical Office of the European Union, 2019

# Data

```{r, echo=FALSE, fig.fullwidth=TRUE, layout="l-page"}
library(DT)

data %>% 
  select(date, age_group, incidence, incidence_7day, incidence_per_capita, incidence_7day_per_capita) %>%
  DT::datatable(extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('csv')
  ))
```

### Imprint and privacy policy

[Imprint and privacy policy](https://covid19-r.com/privacy.html)

This report, the results and figures, as well as the data are made available under the
[Creative Commons Attribution 4.0 International Public License](https://creativecommons.org/licenses/by/4.0/).

Usage of the data requires attribution to the sources as described in the sources section. 

The source code is available on GitHub under 
[fvalka/covid-austria-age-heterogenity](https://github.com/fvalka/covid-austria-age-heterogenity)

